---
title: "homework3_scratchwork"
format: html
---

## 0. Null model
```{r null-model0}
model0 <- lm(total_g ~ 1, # formula
             data = drought_exp_clean) # data frame
```


## 1. total biomass as a function of SLA, water treatment, and species
```{r saturated-model1}
# saturated model
model1 <- lm(total_g ~ sla + water_treatment + species_name,
             data = drought_exp_clean)
```

## 2. total biomass as a function of SLA and water treatment
```{r simpler-model2}
model2 <- lm(total_g ~ sla + water_treatment,
             data = drought_exp_clean)
```

## 3. total biomass as a function of SLA and species
```{r simpler-model3}
model3 <- lm(total_g ~ sla + species_name,
             data = drought_exp_clean)

```

## 4. MY MODEL (last one)
```{r simpler-model4}
model4 <- lm(total_g ~ water_treatment + species_name,
             data = drought_exp_clean)

summary(model4)

par(mfrow = c(2, 2))
plot(model4) 
```




```{r model-predictions-with-ggplot}
# creating new data frame of model predictions for plotting
model_preds_for_plotting <- model_preds %>% 
  rename(sla = x, # renaming columns to make this easier to use
         water_treatment = group,
         species_name = facet)

# use View(model_preds_for_plotting) 
# to compare this to the original model_preds data frame

ggplot() +
  # underlying data
  geom_point(data = drought_exp_clean,
             aes(x = sla,
                 y = total_g,
                 color = water_treatment)) +
  # model prediction 95% CI ribbon
  geom_ribbon(data = model_preds_for_plotting,
              aes(x = sla, 
                  y = predicted,
                  ymin = conf.low,
                  ymax = conf.high,
                  fill = water_treatment),
              alpha = 0.2) +
  # model prediction lines
  geom_line(data = model_preds_for_plotting,
            aes(x = sla, 
                y = predicted,
                color = water_treatment)) +
  # cleaner theme
  theme_classic() +
  # creating different panels for species
  facet_wrap(~species_name) 
```


```{r model-selection}
library(MuMIn)

model.sel(model0,
          model1, 
          model2, 
          model3, 
          model4)
```

```{r model-tables}
# one option for a single model
flextable::as_flextable(model4) 

# another option for a single model
gtsummary::tbl_regression(model1)

# comparing models
modelsummary::modelsummary( # this function takes a list of models
  list( 
    "null" = model0, # "model name" = model object
    "model 1" = model1,
    "model 2" = model2,
    "model 3" = model3, 
    "model 4" = model4
  )
) 


Table: In your table, each row should be a model with the model number (1, 2, 3, etc.) and the predictors for each model.
```


b. Write a 5-6 sentence “statistical methods” section. (8 points)
Your answer should be in paragraph form and include:

how you addressed the central question(s) (i.e. to examine the influence of ____, ____, and ____ on _____, I…)

I ran 5 different statistical models (null, etc) and compared using AIc and delta thing to identify best model, used multiple lin reg model, categorical var, 

how you chose the final model (i.e. to determine the model that best described ____, I…)

AIc and delta 

how you visually determined that your final model conformed to the assumptions of a linear model (i.e. to evaluate linear model assumptions, I…)

see outputs
par(mfrow = c(2, 2))
plot(model4)  







last part cite  p-value: < 2.2e-16 and Adjusted R-squared:  0.7257 and the AIC/delta



education psychology 







