---
title: "Lund-Paige_homework-03"
format: 
  html: 
    toc: true #include table of contents in rendered html 
execute: #all code chunks
    warning: false #no warnings
    message: false #no messages
---

Link  to forked repository: https://github.com/pllund/Lund-Paige_homework-03 Note: document should be committed and pushed ten times

# Problem 1: Multiple Linear Regression

## Setting Up
```{r set-up}
#load in necessary general packages
library(tidyverse)
library(readxl)
library(here)
library(janitor)
#package for visualizing pairs
library(GGally)
#package for model selection
library(MuMIn)
#package for model predictions
library(ggeffects)
#package for model tables
library(gtsummary)
library(flextable)
library(modelsummary)

#load in drought data 
drought_exp <- read_xlsx(path = here("data", 
                                     "Valliere_etal_EcoApps_Data.xlsx"),
                         sheet = "First Harvest")

#look at data to check it is correct
str(drought_exp)
class(drought_exp)
```
## Data Cleaning
```{r cleaning}
#clean the drought_exp data
drought_exp_clean <- drought_exp %>% 
  #set nicer column names
  clean_names() %>% 
  #add column with species scientific names
  mutate(species_name = case_when( 
    species == "ENCCAL" ~ "Encelia californica", # bush sunflower
    species == "ESCCAL" ~ "Eschscholzia californica", # California poppy
    species == "PENCEN" ~ "Penstemon centranthifolius", # Scarlet bugler
    species == "GRICAM" ~ "Grindelia camporum", # great valley gumweed
    species == "SALLEU" ~ "Salvia leucophylla", # Purple sage
    species == "STIPUL" ~ "Nasella pulchra", # Purple needlegrass
    species == "LOTSCO" ~ "Acmispon glaber" # deerweed
  )) %>% 
  #moving species_name column after species
  relocate(species_name, .after = species) %>% 
  #adding column with full treatment names
  mutate(water_treatment = case_when( 
    water == "WW" ~ "Well watered",
    water == "DS" ~ "Drought stressed"
  )) %>% 
  #moving water_treatment column after water
  relocate(water_treatment, .after = water) 
```

## Visualizing Data

### Correlations
```{r}


```


# Problem 2: Affective Visualization

a. Describe in words what an affective visualization could look like for your personal data (3-5 sentences). 




# Problem 3: Statistical Critique

a. Revisit and Summarize: What are the statistical tests the authors are using to address their main research question? Insert the figure or table you described in Homework 2 here.

**The author used a logistic regression statistical test to answer their main research question which was identifying whether there was a relationship between tobacco outlet density near schools and individual's smoking habits. Additionally, they used a linear regression model to test the relationship between density, perceived prevalence and peer approval.**

```{r}
#load packages for importing png and set directory
library(png)
library(grid)
#setwd("/Users/paigelund/desktop/es_193DS")
#read in png
#table <- readPNG("table.png")
#show png
#grid.raster(table)
```








